---
title: "BDA - Assignment 4"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 3
---


```{r setup, include=FALSE}
# This chunk just sets echo = TRUE as default (i.e. print all code)
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=TRUE, include=FALSE}
# To install aaltobda, see the General information in the assignment.
library(aaltobda)
library(mvtnorm)
library(markmyassignment)
library(ggplot2)
#exercise_path <-
#"https://github.com/avehtari/BDA_course_Aalto/blob/master/exercises/tests/ex4.yml"
#set_assignment(exercise_path)
#mark_my_assignment()
```


## Exercise 1. 
### a)
Provided that $corr(\alpha,\beta)=\frac{Covariance(\alpha,\beta)}{\sqrt{Variance(\alpha)Variance(\beta)}}$, the formula of the covariance matrix $$\left[\begin{matrix}
 Cov(\alpha,\alpha)& Cov(\alpha,\beta)\\ 
 Cov(\beta,\alpha)& Cov(\beta,\beta)
\end{matrix}\right]$$ and also $Cov(\alpha,\alpha)=2^2, Cov(\beta,\beta)=10^2$, we can get \textbf{the mean and the covariance of the bivariate distribution} $$\mu=[0,10]$$ and $$\Sigma=\left[\begin{matrix}
4& 10\\ 
10& 100
\end{matrix}\right]$$.

### b)
The code for the function that computes the logarithm of the density of the prior distribution in a) for arbitrary values of $\alpha$ and $\beta$ goes as follows.

```{r}
p_log_prior<-function(alpha, beta){
return(log(dmvnorm(c(alpha,beta),c(0,10),matrix(c(4,10,10,100),ncol=2,nrow=2))))
}
```
Also we can have a check here using the data from the instruction. The function should return a value very close to the correct answer.
```{r}
# have a test
alpha <- 3
beta <- 9
cat("the correct answer is:  -6.296435 \n")
cat("the calculated answer is: ",p_log_prior(alpha,beta),"\n")
cat("the difference is",p_log_prior(alpha,beta)+6.296435,"\n")
```


### c)
The code for the function that computes the logarithm of the density of the posterior for arbitrary values of $\alpha$ and $\beta$ goes as follows.
```{r}
data("bioassay")
p_log_posterior<-function(alpha, beta){
prior<-log(dmvnorm(c(alpha,beta),c(0,10),matrix(c(4,10,10,100),ncol=2,nrow=2)))
likelihood<-bioassaylp(alpha,beta,bioassay$x,bioassay$y,bioassay$n)
return(prior+likelihood)
}
```
Also we can have a check here using the data from the instruction. The function should return a value very close to the correct answer.
```{r}
# have a test
alpha <- 3
beta <- 9
cat("the correct answer is:  -15.78798 \n")
cat("the calculated answer is: ",p_log_posterior(alpha,beta),"\n")
cat("the difference is",p_log_posterior(alpha,beta)+15.78798,"\n")
```

### d)

The figure is shown below.

```{r}
bioassay_posterior_density_plot(alpha_limits=c(-4,4),beta_limits = c(-10,30))
```

### e)

The code for the functions that computes the importance ratios and the normalized weights for each sample goes as follows.

```{r}
log_importance_weights<-function(alpha, beta){
likelihoods=c()
for (i in seq(1,length(alpha),length=length(alpha)))
{
likelihoods=c(likelihoods,bioassaylp(alpha[i],beta[i],bioassay$x,bioassay$y,bioassay$n))
}
return(likelihoods)
}

normalized_importance_weights<-function(alpha, beta)
{
weights<-log_importance_weights(alpha,beta)
weights<-exp(weights)
nor_weights<-weights/sum(weights)
return(nor_weights)
}
```
Also we can have a check here using the data from the instruction. The function should return a value very close to the correct answer.
```{r}
#have a test
alpha <- c(1.896, -3.6, 0.374, 0.964, -3.123, -1.581)
beta <- c(24.76, 20.04, 6.15, 18.65, 8.16, 17.4)
cat("Here's the test case:")
cat("the correct answer for the log weights is: -8.95 -23.47 -6.02 -8.13 -16.61 -14.57 \n")
cat("the calculated answer for the log weights is: ",round(log_importance_weights(alpha,beta),2),'\n')
cat("the difference is ",mean(log_importance_weights(alpha,beta)-c(-8.95,-23.47,-6.02,-8.13,-16.61,-14.57)),'\n')
cat("the correct answer for the normalized weights is:  0.045 0.000 0.852 0.103 0.000 0.000 \n")
cat("the calculated answer for the normalized weights is: ",round(normalized_importance_weights(alpha,beta),3),'\n')
cat("the difference is ",mean(normalized_importance_weights(alpha,beta)-c(0.045,0.000,0.852,0.103,0.000,0.000)),'\n')
cat("test case ends")
```
Now that all the functions defined above work correctly. Now we can forget about the test case and draw some real samples from the prior.
```{r}
samples<-rmvnorm(10000,c(0,10),matrix(c(4,10,10,100),ncol=2,nrow=2))   #draw some samples
alpha_samples<-samples[,1]
beta_samples<-samples[,2]
cat("real case:10000 samples drawn")
```

### f)

```{r}
posterior_mean<-function(alpha, beta)
{
weights<-normalized_importance_weights(alpha,beta)
alpha_mean<-sum(weights*alpha)/sum(weights)
beta_mean<-sum(weights*beta)/sum(weights)
return(c(alpha_mean,beta_mean))
}

cat("the posterior mean using importance sampling and draws from e) is: ",
    posterior_mean(alpha_samples,beta_samples))
```


### g)

```{r}
S_eff<-function(alpha, beta)
{
weights<-normalized_importance_weights(alpha,beta)
return(1/sum(weights^2))
}
cat("the effective sample size is: ",S_eff(alpha_samples,beta_samples))
```

### h)

The scatterplot of the obtained posterior sample is shown below.

```{r}
samples<-rmvnorm(10000,c(0,10),matrix(c(4,10,10,100),ncol=2,nrow=2))
alpha<-samples[,1]
beta<-samples[,2]
weights<-normalized_importance_weights(alpha,beta)
alpha_resample<-sample(alpha, 1000, replace = FALSE, prob = weights)
beta_resample<-sample(beta, 1000, replace = FALSE, prob = weights)

plot(alpha_resample,beta_resample,'p', xlab="Alpha Samples", ylab="Beta Samples",
   main="Posterior Scatter Plot")
```



### i)

```{r}
cat("the estimate for p(beta > 0 | x,n,y) (probability that the drug is harmful)  is:",length(which(beta_resample>0))/length(beta_resample),"\n")
```


### j)
The histogram of the draws from the posterior distribution of the LD50 conditional on Beta > 0 is shown below
```{r}
#delete points with beta<0
alpha_resample_conditional<-c()
beta_resample_conditional<-c()
for (i in seq(1,1000,len=1000))
{
  if (beta_resample[i]>0)
  {
    alpha_resample_conditional<-c(alpha_resample_conditional,alpha_resample[i])
    beta_resample_conditional<-c(beta_resample_conditional,beta_resample[i])
  }
}
hist(-alpha_resample_conditional/beta_resample_conditional,xlab="LD50",main="histogram of LD50 conditional on Beta > 0")
ggplot() + geom_histogram(aes(-alpha_resample_conditional/beta_resample_conditional))+
  labs(title="histogram of LD50 conditional on Beta > 0",x="LD50", y = "Count")
```