---
title: "BDA - Assignment 8"
output:
  pdf_document: 
    toc: yes
    toc_depth: 2
---


```{r setup, include=FALSE}
# This chunk just sets echo = TRUE as default (i.e. print all code)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(aaltobda)
library(ggplot2)
library(StanHeaders)
library(rstan)
library(loo)
data("factory")
```
# 1) Pooled Model
In the pooled model all the machines are considered as one entity, thus we have to combine the all the measurements into one and perform our prediction on the whole data, rather than a subset.   
The model is defined as follows. 
```{r,eval=FALSE}
data {
int<lower=0> N; // number of data points
vector[N] y; //
}
parameters {
real mu; // group means
real<lower=0> sigma; // common std
}
model {
y ~ normal(mu, sigma);
}

generated quantities {
real ypred;
ypred = normal_rng(mu, sigma);
vector[N] log_lik;
for (i in 1:N)
log_lik[i] = normal_lpdf(y[i] | mu, sigma);
}

```
Then we need to fit it and use it to generate desired values. 
```{r, message=FALSE,results="hide"}
pooled_factory_data<-as.vector(as.matrix(factory))
pooled_data<-list(N=length(pooled_factory_data),
                  y=pooled_factory_data)
pooled_fit<-stan(file='pooled_model_assignment8.stan',data=pooled_data)
```
```{r, warning=FALSE}
pooled_fit_ss <- extract(pooled_fit, permuted = TRUE) 
pooled_log_lik <- pooled_fit_ss$log_lik
dims_pooled<-dim(pooled_log_lik)
S<-dims_pooled[1]
pooled_psis<-loo(pooled_log_lik)
pooled_elpd_loo<-pooled_psis$estimates[1,1]
pooled_p_loo<-pooled_psis$estimates[2,1]
cat("The PSIS-LOO value for pool model is:",pooled_elpd_loo,'\n' )
cat("The p_ef value for pool model is:",pooled_p_loo ,'\n')
pooled_k_values<-pooled_psis$diagnostics$pareto_k
hist(pooled_k_values,main ="K of PSIS-LOO with pool model",xlim=c(0.0,1.0))
```


# 2) Separate Model



In the separate model we treat every machine as an individual entity, thus when calculating $\sigma$ or $\mu$ we take into consideration only a single machine. The combination of all machines should not effect $\sigma$ or $\mu$.
The stan model is defined as follows.
```{r,eval=FALSE}
data {
int<lower=0> N; // number of data points
int<lower=0> K; // number of groups
int<lower=1,upper=K> x[N]; // group indicator
vector[N] y;
}
parameters {
vector[K] mu; // group means
vector<lower=0>[K] sigma; // group stds
}
model {
y ~ normal(mu[x], sigma[x]);
}
generated quantities {
real ypred;
vector[N] log_lik;
ypred = normal_rng(mu[6], sigma[6]);
for (i in 1:N)
    log_lik[i] = normal_lpdf(y[i] | mu[x[i]], sigma[x[i]]);
}
```
Then we need to fit it and use it to generate desired values. 
```{r, message=FALSE,results="hide"}
separate_data<-list(N=length(as.matrix(factory)),
                    K=6,
                    x=c(1, 1, 1, 1, 1,2, 2, 2, 2, 2,3, 3, 
                        3, 3, 3,4, 4, 4, 4, 4,5, 5, 5, 5, 5,6, 6, 6, 6, 6),
                    y=as.vector(as.matrix(factory)))
separate_fit<-stan(file='separate_model_assignment8.stan',data=separate_data)
```

```{r, warning=FALSE}
separate_fit_ss <- extract(separate_fit, permuted = TRUE) 
separate_log_lik <- separate_fit_ss$log_lik
dims_separate<-dim(separate_log_lik)
S<-dims_separate[1]
separate_psis<-loo(separate_log_lik)
separate_elpd_loo<-separate_psis$estimates[1,1]
separate_p_loo<-separate_psis$estimates[2,1]
cat("The PSIS-LOO value for separate model is:",separate_elpd_loo,'\n' )
cat("The p_ef value for separate model is:",separate_p_loo ,'\n')
separate_k_values<-separate_psis$diagnostics$pareto_k
hist(separate_k_values,main ="K of PSIS-LOO with separate model",xlim=c(0.0,1.0))

separate_elpd_loo<-separate_psis$estimates[1,1]
separate_p_loo<-separate_psis$estimates[2,1]




```




# 3) Hierarchical Model
The hierarchical model is quite interesting in the sense that it can predict measurements for the machines which have no data. For example, there is no data about the seventh machine, but this model can predict its posterior distribution.  
The stan model is defined as follows.
```{r,eval=FALSE}
data {
int<lower=0> N; // number of data points
int<lower=0> K; // number of groups
int<lower=1,upper=K> x[N]; // group indicator
vector[N] y;
}
parameters {
real mu0; // prior mean
real<lower=0> sigma0; // prior std
vector[K] mu; // group means
real<lower=0> sigma; // common std
}
model {
mu ~ normal(mu0, sigma0);
y ~ normal(mu[x], sigma);
}
generated quantities {
real ypred6;
real mu7;
vector[N] log_lik;
ypred6 = normal_rng(mu[6], sigma);
mu7 = normal_rng(mu0, sigma0);
for (i in 1:N)
    log_lik[i] = normal_lpdf(y[i] | mu[x[i]], sigma);
}
```
Then we need to fit it and use it to generate desired values.    
```{r, message=FALSE,results="hide"}
hierarchical_fit<-stan(file='hierarchical_model_assignment8.stan',data=separate_data)
```

```{r, warning=FALSE}
hierarchical_fit_ss <- extract(hierarchical_fit, permuted = TRUE) 
hierarchical_log_lik <- hierarchical_fit_ss$log_lik
dims_hierarchical<-dim(hierarchical_log_lik)
S<-dims_hierarchical[1]
hierarchical_psis<-loo(hierarchical_log_lik)
hierarchical_elpd_loo<-hierarchical_psis$estimates[1,1]
hierarchical_p_loo<-hierarchical_psis$estimates[2,1]
cat("The PSIS-LOO value for hierarchical model is:",hierarchical_elpd_loo,'\n' )
cat("The p_ef value for hierarchical model is:",hierarchical_p_loo ,'\n')
hierarchical_k_values<-hierarchical_psis$diagnostics$pareto_k
hist(hierarchical_k_values,main ="K of PSIS-LOO with hierarchical model",xlim=c(0.0,1.0))
```
# 4) Summary
```{r}
cat("The PSIS-LOO value for pool model is:",pooled_elpd_loo,'\n' )
cat("The p_ef value for pool model is:",pooled_p_loo ,'\n')
cat("The PSIS-LOO value for separate model is:",separate_elpd_loo,'\n' )
cat("The p_ef value for separate model is:",separate_p_loo ,'\n')
cat("The PSIS-LOO value for hierarchical model is:",hierarchical_elpd_loo,'\n' )
cat("The p_ef value for hierarchical model is:",hierarchical_p_loo ,'\n')
```
Based on the histograms of $\hat{k}$ values of these three models, we can see that the PSIS-LOO estimates of pool model is the most reliable one since its $\hat{k}$ values are all lower than 0.7. And the PSIS-LOO estimates of separate model is the least reliable one. For the PSIS-LOO estimates of hierarchical model, most of its $\hat{k}$ values are lower than 0.7 so it's totally acceptable. 

Thus, according to PSIS-LOO, we should choose hierarchical model. 

















